---
title: "Openpayments data analysis"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    # theme: yeti
    code_folding: hide
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(lubridate)
library(writexl)
options(dplyr.summarise.inform = FALSE)

#target drug query done in BigQuery. Resulting data slice exported to csv and we'll load them here:
targetdrugsonly <- read_csv("processed_data/targetdrugsonly.csv", col_types = cols(.default = "c"))

glimpse(targetdrugsonly)

#clean up column names
targetdrugsonly <- targetdrugsonly %>% 
  clean_names()

#format columns
targetdrugsonly <- targetdrugsonly %>% 
  mutate(
    total_amount_of_payment_us_dollars = as.numeric(total_amount_of_payment_us_dollars),
    date_of_payment = ymd(date_of_payment),
    number_of_payments_included_in_total_amount = as.numeric(number_of_payments_included_in_total_amount),
    payment_publication_date = ymd(payment_publication_date)
  ) 

glimpse(targetdrugsonly)

#if there are any case issues with the doctor names or drug names, let's standardize that
targetdrugsonly <- targetdrugsonly %>% 
  mutate(
    physician_first_name = str_trim(str_to_upper(physician_first_name)),
    physician_middle_name = str_trim(str_to_upper(physician_middle_name)),
    physician_last_name = str_trim(str_to_upper(physician_last_name)),
    recipient_city = str_trim(str_to_upper(recipient_city)),
    name_of_drug_or_biological_or_device_or_medical_supply_1 = str_trim(str_to_upper(name_of_drug_or_biological_or_device_or_medical_supply_1)),
    name_of_drug_or_biological_or_device_or_medical_supply_2 = str_trim(str_to_upper(name_of_drug_or_biological_or_device_or_medical_supply_2)),
    name_of_drug_or_biological_or_device_or_medical_supply_3 = str_trim(str_to_upper(name_of_drug_or_biological_or_device_or_medical_supply_3)),
    name_of_drug_or_biological_or_device_or_medical_supply_4 = str_trim(str_to_upper(name_of_drug_or_biological_or_device_or_medical_supply_4)),
    name_of_drug_or_biological_or_device_or_medical_supply_5 = str_trim(str_to_upper(name_of_drug_or_biological_or_device_or_medical_supply_5))
  )


#since the drug names are split across five columns, we'll make a combined one to make searching easier
targetdrugsonly <- targetdrugsonly %>% 
  mutate(
    ak_drugname_combined = paste(name_of_drug_or_biological_or_device_or_medical_supply_1,
                              name_of_drug_or_biological_or_device_or_medical_supply_2,
                              name_of_drug_or_biological_or_device_or_medical_supply_3,
                              name_of_drug_or_biological_or_device_or_medical_supply_4,
                              name_of_drug_or_biological_or_device_or_medical_supply_5,
                              sep = "|")
  )
    


```

Looking at the most recent four years of doctor-payments data from CMS (2016-2019) for the target drugs in question.

## Quick Aggregates

#### Total money overall 
```{r}
#total money?
targetdrugsonly %>% 
  summarise(totalpayments = sum(total_amount_of_payment_us_dollars))
```

#### Total by year
```{r}
targetdrugsonly %>% 
  group_by(program_year) %>% 
  summarise(totalpayments = sum(total_amount_of_payment_us_dollars))
```

#### Total by recipient type
Some went to teaching hospitals, but most to doctors
```{r}
targetdrugsonly %>% 
  group_by(covered_recipient_type) %>% 
  summarise(totalpayments = sum(total_amount_of_payment_us_dollars))
```

#### By company who submitted the records
```{r}
#by company
targetdrugsonly %>% 
  group_by(submitting_applicable_manufacturer_or_applicable_gpo_name) %>% 
  summarise(totalpayments = sum(total_amount_of_payment_us_dollars)) %>% 
  arrange(desc(totalpayments))
```

#### By company who gave the payments
```{r}
targetdrugsonly %>% 
  group_by(applicable_manufacturer_or_applicable_gpo_making_payment_name) %>% 
  summarise(totalpayments = sum(total_amount_of_payment_us_dollars)) %>% 
  arrange(desc(totalpayments))
```

#### Top 10 drug combos
Since a payment can be associated with up to five different drugs, we've combined those five
into a single column to make it a bit easier to work with. Here are the top combinations
to provide a sense of which drugs may be responsible of the most payments.
```{r}
#drug combo variations
targetdrugsonly %>% 
  group_by(ak_drugname_combined) %>% 
  summarise(totalpayments = sum(total_amount_of_payment_us_dollars)) %>% 
  arrange(desc(totalpayments)) %>% 
  head(10)
```

#### Top 10 doctors on the receiving end
```{r}
#top docs
targetdrugsonly %>% 
  group_by(physician_last_name, physician_first_name, recipient_state) %>% 
  summarise(totalpayments = sum(total_amount_of_payment_us_dollars)) %>%
  arrange(desc(totalpayments)) %>% 
  head(10)
```
#### Type of payments/gifts
```{r}
#top docs
targetdrugsonly %>% 
  # filter(covered_recipient_type == "Covered Recipient Physician") %>% 
  group_by(nature_of_payment_or_transfer_of_value) %>% 
  summarise(totalpayments = sum(total_amount_of_payment_us_dollars)) %>%
  arrange(desc(totalpayments)) 
```
#### Country of travel
```{r}
#top docs
targetdrugsonly %>% 
  filter(nature_of_payment_or_transfer_of_value == "Travel and Lodging") %>% 
  group_by(country_of_travel) %>% 
  summarise(totalpayments = sum(total_amount_of_payment_us_dollars)) %>%
  arrange(desc(totalpayments)) 

```


#### City of travel
```{r}
#top docs
targetdrugsonly %>% 
  filter(nature_of_payment_or_transfer_of_value == "Travel and Lodging") %>% 
  group_by(country_of_travel, city_of_travel) %>% 
  summarise(totalpayments = sum(total_amount_of_payment_us_dollars)) %>%
  arrange(desc(totalpayments)) %>% 
  head(10)

```


